---
title: "Data Analysis Assignment #2 (75 points total)"
author: "Venit_Michael"
output:
  html_document: default
---

```{r setup, include = FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

```


##Data Analysis #2

```{r analysis_setup1, message = FALSE, warning = FALSE}

# Perform the following steps to start the assignment.
 
# 1) Load/attach the following packages via library():  flux, ggplot2, gridExtra, moments, rockchalk, car.
# NOTE:  packages must be installed via install.packages() before they can be loaded.

library(dplyr)
library(flux)
library(ggplot2)
library(gridExtra)
library(knitr)
library(rockchalk)
library(tidyverse)

# 2) Use the "mydata.csv" file from Assignment #1 or use the file posted on the course site.  Reading
# the files into R will require sep = "" or sep = " " to format data properly.  Use str() to check file
# structure.

mydata <- read.csv("~/Desktop/mydata.csv", sep = ",", stringsAsFactors = TRUE)

# mydata <- read.csv(file.path("c:...", "mydata.csv"), sep = ",")
# mydata <- read.csv(file.path("c:/Rabalone/", "mydata.csv"), sep = ",")

str(mydata)

```

### Test Items starts from here - There are 10 sections - total of 75 points ##############

***#### Section 1: (5 points) ####***

(1)(a) Form a histogram and QQ plot using RATIO. Calculate skewness and kurtosis using 'rockchalk.' Be aware that with 'rockchalk', the kurtosis value has 3.0 subtracted from it which differs from the 'moments' package. 

```{r Part_1a, fig.width = 12}
skew <- skewness(mydata$RATIO)
kurt <- kurtosis(mydata$RATIO)


rock_skew <- rockchalk::skewness(mydata$RATIO)
rock_kurt <- rockchalk::kurtosis(mydata$RATIO, excess = FALSE)

par(mfrow = c(1,2))

hist(mydata$RATIO, col = "dark green", xlab = "Ratio", main = "Shuck/Volum Ratio")

qqnorm(mydata$RATIO, main = "QQ Plot", col = ifelse(mydata$RATIO %in% c(boxplot.stats(mydata$RATIO)$out), "dark green", "red"))
qqline(mydata$RATIO)

cat("Skewness (base R)    : ", skew, "\n")
cat("Skewness (rockchalk) : ", rock_skew, "\n")
cat("Kurtosis (base R)    : ", kurt, "\n")
cat("Kurtosis (rockchalk) : ", rock_kurt, "\n")

```

(1)(b) Tranform RATIO using *log10()* to create L_RATIO (Kabacoff Section 8.5.2, p. 199-200). Form a histogram and QQ plot using L_RATIO. Calculate the skewness and kurtosis. Create a boxplot of L_RATIO differentiated by CLASS.

```{r Part_1b, fig.width = 12, fig.height = 8}

mydata$L_RATIO <- log10(mydata$RATIO)
cat("Using rockchalk to calculate skewness and kurtosis [L_RATIO]", "\n\n")
rock_skew3 <- rockchalk::skewness(mydata$L_RATIO)
rock_kurt3 <- rockchalk::kurtosis(mydata$L_RATIO, excess = FALSE)

cat("Skewness: ", rock_skew3, "\n")
cat("Kurtosis: ", rock_kurt3, "\n")

par(mfrow= c(1,2))

hist(mydata$L_RATIO, col= "dark green", xlab = "Log(Ratio)", main= "Histogram Log(Ratio)")

qqnorm(mydata$L_RATIO, main = "QQ Plot: Log(Ratio)", col= ifelse(mydata$L_RATIO %in% c(boxplot.stats(mydata$L_RATIO)$out), "dark green", "red"))

qqline(mydata$L_RATIO)

boxplot(mydata$L_RATIO ~ mydata$CLASS, main= "Log(Ratio) by Class", ylab = "Log(Ratio)", xlab = "Class", beside= TRUE, col= "dark green")

```

(1)(c) Test the homogeneity of variance across classes using *bartlett.test()* (Kabacoff Section 9.2.2, p. 222). 

```{r Part_1c}
ratio<- bartlett.test(RATIO ~ CLASS, data= mydata)
Lratio<- bartlett.test(L_RATIO ~ CLASS, data= mydata)

ratio
Lratio
```

**Essay Question: Based on steps 1.a, 1.b and 1.c, which variable RATIO or L_RATIO exhibits better conformance to a normal distribution with homogeneous variances across age classes?  Why?** 

***Answer: (Enter your answer here.)***
Log(ratio) by class shows a smaller Bartlett K-squared compared to ratio by class. Log(ratio) exhibits better conformance to a normal distribution with homogenous variances across age classes as the histogram proves to be less skewed with reduced kurtosis when compared to the histogram that compares ratio by class. 

***#### Section 2 (10 points) ####***

(2)(a) Perform an analysis of variance with *aov()* on L_RATIO using CLASS and SEX as the independent variables (Kabacoff chapter 9, p. 212-229). Assume equal variances. Perform two analyses. First, fit a model with the interaction term CLASS:SEX. Then, fit a model without CLASS:SEX. Use *summary()* to obtain the analysis of variance tables (Kabacoff chapter 9, p. 227).

```{r Part_2a}
summary(aov(L_RATIO ~ CLASS + SEX + CLASS:SEX, data= mydata))
summary(aov(L_RATIO ~ CLASS + SEX, data= mydata))
```

**Essay Question:  Compare the two analyses.  What does the non-significant interaction term suggest about the relationship between L_RATIO and the factors CLASS and SEX?**

***Answer: (Enter your answer here.)***
CLASS:SEX does not contain a significant p-value when predicting L_RATIO compared to CLASS and SEX. The independent variable had little effect and was deemed insignificant. However, CLASS and SEX, are significant when predicting L_RATIO. 

(2)(b) For the model without CLASS:SEX (i.e. an interaction term), obtain multiple comparisons with the *TukeyHSD()* function. Interpret the results at the 95% confidence level (*TukeyHSD()* will adjust for unequal sample sizes). 

```{r Part_2b}

TukeyHSD(aov(L_RATIO ~ CLASS + SEX, data = mydata))

```

**Additional Essay Question:  first, interpret the trend in coefficients across age classes. What is this indicating about L_RATIO?  Second, do these results suggest male and female abalones can be combined into a single category labeled as 'adults?' If not, why not?**

***Answer: (Enter your answer here.)***
The pattern of coefficients as it pertains to L_RATIO by Class is mostly negative. It is safe to assume that there is a uniform, negative correlation between the two afformentioned variables. 

There does not appear to be a significant difference between classes A1 and A2, due to the high p value = 0.6919456. However, there are notable significant differences between all other classes; all other class comparisons exceed an the arbitrary p-value threshold of 0.05.   

Based on the high p-value of 0.9412689 between male and female abalones, it is plausible to conclude that there is not a significant difference between male and female abalones. No arbitrary p-value threshold would allow us to reject this null hypothesis. Combining sexes into a single category, "adults", might prove to be more beneficial. 


***####  Section 3: (10 points) ####***

(3)(a1) Here, we will combine "M" and "F" into a new level, "ADULT". The code for doing this is given to you. For (3)(a1), all you need to do is execute the code as given.

```{r Part_3a1}

mydata$TYPE <- factor(ifelse(mydata$SEX == "I", "I", "ADULT"))
table(mydata$TYPE)


```

(3)(a2)  Present side-by-side histograms of VOLUME. One should display infant volumes and, the other, adult volumes. 

```{r Part_3a2, fig.width = 12}
par(mfrow= c(1,2))

hist(mydata$VOLUME[mydata$TYPE== "I"], col = "dark blue", xlab = "Volume", main = "Volume by Infants", ylim = c(0,150))

hist(mydata$VOLUME[mydata$TYPE== "ADULT"], col = "dark green", xlab = "Volume", main = "Volume by Infants", ylim = c(0,150))
```


**Essay Question: Compare the histograms.  How do the distributions differ? Are there going to be any difficulties separating infants from adults based on VOLUME?**

***Answer: (Enter your answer here.)***
The distributions differ drastically. When observing volume by infants, the distribution is skewed to the right. The mode for the afformentioned histogram is 200, but there do appear to be some instances with higher volume over 600. As it pertains to the histogram that portrays volume by adults, the distribution appears more normal. While the mode is between 300-500. It does not appear that separating adults and infants by volume is advantageous for the sake od this study as can be observed by the differing means in both groups. 

(3)(b) Create a scatterplot of SHUCK versus VOLUME and a scatterplot of their base ten logarithms, labeling the variables as L_SHUCK and L_VOLUME. Please be aware the variables, L_SHUCK and L_VOLUME, present the data as orders of magnitude (i.e. VOLUME = 100 = 10^2 becomes L_VOLUME = 2). Use color to differentiate CLASS in the plots. Repeat using color to differentiate by TYPE. 

```{r Part_3b, fig.width = 12, fig.height = 8}
mydata$L_SHUCK<- log10(mydata$SHUCK)
mydata$L_VOL<- log10(mydata$VOLUME)

par(mfrow= c(2,4))

plot(mydata$VOLUME, mydata$SHUCK, col= mydata$CLASS, main = "Shuck v. Volume by Class", xlab = "Volume", ylab = "Shuck")
legend('bottomright', legend = levels(mydata$CLASS), bty = "n", col = 1:5, cex = 0.75, pch = 1)

plot(mydata$L_VOL, mydata$L_SHUCK, col= mydata$CLASS, main = "L_Shuck v. L_Volume by Class", xlab = "Volume", ylab = "Shuck")
legend('bottomright', legend = levels(mydata$CLASS), bty = "n", col = 1:5, cex = 0.75, pch = 1)

plot(mydata$VOLUME, mydata$SHUCK, col= mydata$TYPE, main = "Shuck v. Volume by Type", xlab = "Volume", ylab = "Shuck")
legend('bottomright', legend = levels(mydata$TYPE), bty = "n", col = 1:5, cex = 0.75, pch = 1)

plot(mydata$L_VOL, mydata$L_SHUCK, col= mydata$TYPE, main = "L_Shuck v. L_Volume by Type", xlab = "Volume", ylab = "Shuck")
legend('bottomright', legend = levels(mydata$TYPE), bty = "n", col = 1:5, cex = 0.75, pch = 1)

```

**Additional Essay Question:  Compare the two scatterplots. What effect(s) does log-transformation appear to have on the variability present in the plot?  What are the implications for linear regression analysis? Where do the various CLASS levels appear in the plots? Where do the levels of TYPE appear in the plots?**

***Answer: (Enter your answer here.)***
The effects of the log-transformation appear in the form of reduced variability among the variables Volume and Shuck. The form of the plot appears more linear when the logarithmic functions are enacted upon those variables while showing a slightly more positive correlation. I would expect there to be a stronger coefficient of determination for L_VOL and L_Shuck. When observing the plot with the log-transformation, Class A5 appears to be concentrated in the top right portion of the scatterplot while Class A1 appears to be more spread out and apparent in the lower left portion of the plot. As it relates to Type, adult abalones are heavily concentrated in the top right portion of the graph while infants reside more towards the bottom left portion of the graph. The differences are noticable when contrasting these variables with and without logarithmic functions enacted on them. 




***####   Section 4: (5 points) ####***

(4)(a1) Since abalone growth slows after class A3, infants in classes A4 and A5 are considered mature and candidates for harvest. You are given code in (4)(a1) to reclassify the infants in classes A4 and A5 as ADULTS. 

```{r Part_4a1}

mydata$TYPE[mydata$CLASS == "A4" | mydata$CLASS == "A5"] <- "ADULT"
table(mydata$TYPE)

```

(4)(a2) Regress L_SHUCK as the dependent variable on L_VOLUME, CLASS and TYPE (Kabacoff Section 8.2.4, p. 178-186, the Data Analysis Video #2 and Black Section 14.2). Use the multiple regression model: L_SHUCK ~ L_VOLUME + CLASS + TYPE. Apply *summary()* to the model object to produce results.

```{r Part_4a2}
model<- lm(L_SHUCK ~ L_VOL + CLASS + TYPE, data = mydata)

summary(model)
```

**Essay Question:  Interpret the trend in CLASS levelcoefficient estimates? (Hint:  this question is not asking if the estimates are statistically significant. It is asking for an interpretation of the pattern in these coefficients, and how this pattern relates to the earlier displays).**

***Answer: (Enter your answer here.)***
There appears to be a negative trend as class progress from A1-15. The higher the class, the more negative the estimate becomes. Due to this observation, it is presumed that L_SHUCK increases in the earlier classes while decreasing as class number climbs. This reflects the observations seen in the scatterplots from the previous question. 


**Additional Essay Question:  Is TYPE an important predictor in this regression? (Hint:  This question is not asking if TYPE is statistically significant, but rather how it compares to the other independent variables in terms of its contribution to predictions of L_SHUCK for harvesting decisions.)  Explain your conclusion.**

***Answer: (Enter your answer here.)***
Type does not appear to be nearly as important of a predictor, as seen from the estimate of -0.021. Type does however seem to have more of an effect on Class A2 when observing magnitudes. Overall, Type will not be of much help when predicting L_SHUCK, rather L_VOL would be a good fit for predicting L_SHUCK. 
-----

The next two analysis steps involve an analysis of the residuals resulting from the regression model in (4)(a) (Kabacoff Section 8.2.4, p. 178-186, the Data Analysis Video #2).

-----

***#### Section 5: (5 points) ####***

(5)(a) If "model" is the regression object, use model$residuals and construct a histogram and QQ plot. Compute the skewness and kurtosis. Be aware that with 'rockchalk,' the kurtosis value has 3.0 subtracted from it which differs from the 'moments' package. 

```{r Part_5am, fig.width = 12}
cat("Using rockchalk to calculate skewness and kurtosis [model$residuals]", "\n\n")

Res_Skew<- rockchalk::skewness(model$residuals)
Res_Kurt<- rockchalk::kurtosis(model$residuals)

cat("Skewness: ", Res_Skew, "\n")
cat("Kurtosis: ", Res_Kurt, "\n")

par(mfrow= c(1,2))

hist(model$residuals, col = "dark green", xlab = "Residuals", main = "Residuals Histogram")
qqnorm(model$residuals, main = "Residuals QQ", col= ifelse(model$residuals %in% c(boxplot.stats(model$residuals)$out), "steelblue", "darkblue"))
qqline(model$residuals)
```

(5)(b) Plot the residuals versus L_VOLUME, coloring the data points by CLASS and, a second time, coloring the data points by TYPE. Keep in mind the y-axis and x-axis may be disproportionate which will amplify the variability in the residuals. Present boxplots of the residuals differentiated by CLASS and TYPE (These four plots can be conveniently presented on one page using *par(mfrow..)* or *grid.arrange()*. Test the homogeneity of variance of the residuals across classes using *bartlett.test()* (Kabacoff Section 9.3.2, p. 222).  

```{r Part_5b, fig.width = 12, fig.height = 8}
mydata$RESIDUALS<- model$residuals

par(mfrow= c(2,4))

plot(mydata$L_VOL, mydata$RESIDUALS, col= mydata$CLASS, main = "L_Volume v. Residuals by Class", xlab = "L_Volume", ylab = "Residuals", xlim = c(0,4), ylim = c(-0.4, 0.4))
legend('bottomright', legend = levels(mydata$CLASS), bty = "n", col = 1:5, cex = 0.75, pch = 1)

plot(mydata$L_VOL, mydata$RESIDUALS, col= mydata$TYPE, main = "L_Volume v. Residuals by Type", xlab = "L_Volume", ylab = "Residuals", xlim = c(0,4), ylim = c(-0.4, 0.4))
legend('bottomright', legend = levels(mydata$TYPE), bty = "n", col = 1:5, cex = 0.75, pch = 1)

boxplot(mydata$RESIDUALS ~ mydata$CLASS, main= "Residuals by Class", ylab = "Residuals", xlab = "Class", beside= TRUE, col= "darkgreen")

boxplot(mydata$RESIDUALS ~ mydata$TYPE, main= "Residuals by TYPE", ylab = "Residuals", xlab = "TYPE", beside= TRUE, col= "darkblue")

bartlett.test(RESIDUALS ~ CLASS, data= mydata)

```

**Essay Question:  What is revealed by the displays and calculations in (5)(a) and (5)(b)? Does the model 'fit'?  Does this analysis indicate that L_VOLUME, and ultimately VOLUME, might be useful for harvesting decisions? Discuss.**  

***Answer: (Enter your answer here.)***
Overall, the observed residuals appear to be evenly spread across y = 0. What is noticable is that more data poitns are observed to be clustered in the right portion of the scatterplot and more disperesed on the left. L_Volume could serve as a useful parameter when making decisions regarding harvesting. However, more alterations could be made to have the output of data produce the residuals in a more evenly distributed manner. 
-----

Harvest Strategy:

There is a tradeoff faced in managing abalone harvest. The infant population must be protected since it represents future harvests. On the other hand, the harvest should be designed to be efficient with a yield to justify the effort. This assignment will use VOLUME to form binary decision rules to guide harvesting. If VOLUME is below a "cutoff" (i.e. a specified volume), that individual will not be harvested. If above, it will be harvested. Different rules are possible.The Management needs to make a decision to implement 1 rule that meets the business goal.

The next steps in the assignment will require consideration of the proportions of infants and adults harvested at different cutoffs. For this, similar "for-loops" will be used to compute the harvest proportions. These loops must use the same values for the constants min.v and delta and use the same statement "for(k in 1:10000)."  Otherwise, the resulting infant and adult proportions cannot be directly compared and plotted as requested. Note the example code supplied below.

-----

***#### Section 6: (5 points) ####***

(6)(a) A series of volumes covering the range from minimum to maximum abalone volume will be used in a "for loop" to determine how the harvest proportions change as the "cutoff" changes. Code for doing this is provided.

```{r Part_6a}

idxi <- mydata$TYPE == "I"
idxa <- mydata$TYPE == "ADULT"

max.v <- max(mydata$VOLUME)
min.v <- min(mydata$VOLUME)
delta <- (max.v - min.v)/10000
prop.infants <- numeric(10000)
prop.adults <- numeric(10000)
volume.value <- numeric(10000)

total.infants <- sum(idxi)  
total.adults <- sum(idxa)

for (k in 1:10000) { 
	value <- min.v + k*delta
	volume.value[k] <- value
	prop.infants[k] <- sum(mydata$VOLUME[idxi] <= value)/total.infants
	prop.adults[k] <-  sum(mydata$VOLUME[idxa] <= value)/total.adults
}

```

(6)(b) Our first "rule" will be protection of all infants. We want to find a volume cutoff that protects all infants, but gives us the largest possible harvest of adults. We can achieve this by using the volume of the largest infant as our cutoff. You are given code below to identify the largest infant VOLUME and to return the proportion of adults harvested by using this cutoff. You will need to modify this latter code to return the proportion of infants harvested using this cutoff. Remember that we will harvest any individual with VOLUME greater than our cutoff.

```{r Part_6b}
# Largest infant volume
(max_inf_vol <- max(mydata$VOLUME[mydata$TYPE == "I"]))  # [1] 526.6383

# Proportion of adults harvested
sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > max_inf_vol) /
  total.adults  # [1] 0.2476573

# Add code to calculate the proportion of infants harvested
sum(mydata$VOLUME[mydata$TYPE == "I"] > max_inf_vol) /
  total.infants

# If we use the largest infant volume, we harvest approximately 24.8% of adults and 0%,
# as expected, of infants.

```

(6)(c) Our next approaches will look at what happens when we use the median infant and adult harvest VOLUMEs. Using the median VOLUMEs as our cutoffs will give us (roughly) 50% harvests. We need to identify the median volumes and calculate the resulting infant and adult harvest proportions for both.

```{r Part_6c}

n.infants<- sum(prop.infants<= 0.5)
u.infants<- sum(prop.infants>0.5)
split.infants<- min.v+ (n.infants + 0.5)*delta
split.infants

n.adults<- sum(prop.adults<= 0.5)
u.adults<- sum(prop.adults>0.5)
split.adults<- min.v+ (n.adults + 0.5)*delta
split.adults

# Add code to determine the median infant volume:
med_inf_vol<- median(mydata$VOLUME[mydata$TYPE=="I"])
med_inf_vol
# Add code to calculate the proportion of infants harvested
sum(split.infants)/ n.infants

# Add code to calculate the proportion of adults harvested
sum(split.adults)/ n.adults

# If we use the median infant volume as our cutoff, we harvest almost 50% of our infants
# and a little more than 93% of our adults.


# Add code to determine the median adult volume:
med_adult_vol<- median(mydata$VOLUME[mydata$TYPE=="ADULT"])
med_adult_vol

# Add code to calculate the proportion of infants harvested
sum(min.v + (u.infants +0.5)*delta) / u.infants

# Add code to calculate the proportion of adults harvested
sum(min.v + (u.adults +0.5)*delta) / u.adults

# If we use the median adult volume as our cutoff, we harvest almost 50% of adults
# and approximately 2.4% of infants.

```

(6)(d) Next, we will create a plot showing the infant conserved proportions (i.e. "not harvested," the prop.infants vector) and the adult conserved proportions (i.e. prop.adults) as functions of volume.value. We will add vertical A-B lines and text annotations for the three (3) "rules" considered, thus far:  "protect all infants," "median infant" and "median adult." Your plot will have two (2) curves - one (1) representing infant and one (1) representing adult proportions as functions of volume.value - and three (3) A-B lines representing the cutoffs determined in (6)(b) and (6)(c).

```{r Part_6d, fig.width = 12, fig.height = 6}
plot(x= volume.value, y=prop.adults, col= "dark green", ylab= "Adult Proportion")
lines(x= volume.value, y= prop.infants, col= "black", xlab= "Volume")
abline(v= median(mydata$VOLUME[mydata$TYPE== "I"]), h= 0.5)
text(133.819, .55, "Median Infant 133.8199")
abline(v= median(mydata$VOLUME[mydata$TYPE== "ADULT"]), h= 0.5)
text(384.5584, .55, "Median Adult 384.5584")
abline(v= max(mydata$VOLUME[mydata$TYPE== "I"]), h= 0.5)
text(526.6836, 0.97, "Max Infant 526.6383")
legend("topright", legend= c("Adult", "Infant"), col= "darkgreen", "black", pch= 1)


```

**Essay Question:  The two 50% "median" values serve a descriptive purpose illustrating the difference between the populations. What do these values suggest regarding possible cutoffs for harvesting?** 

***Answer: (Enter your answer here.)***
There appears to be a good cutoff between the adult and infant volumes where it is possible to to mitigate errors. If one were to harvest at a volume greater than 384.5584, it should be possible to harvest 50% of adults without effecting infants. 

-----


More harvest strategies:

This part will address the determination of a cutoff volume.value corresponding to the observed maximum difference in harvest percentages of adults and infants. In other words, we want to find the volume value such that the vertical distance between the infant curve and the adult curve is maximum. To calculate this result, the vectors of proportions from item (6) must be used. These proportions must be converted from "not harvested" to "harvested" proportions by using (1 - prop.infants) for infants, and (1 - prop.adults) for adults. The reason the proportion for infants drops sooner than adults is that infants are maturing and becoming adults with larger volumes.

-----

***#### Section 7: (10 points)  ####***

(7)(a) Evaluate a plot of the difference ((1 - prop.adults) - (1 - prop.infants)) versus volume.value. Compare to the 50% "split" points determined in (6)(a). There is considerable variability present in the peak area of this plot. The observed "peak" difference may not be the best representation of the data. One solution is to smooth the data to determine a more representative estimate of the maximum difference.

```{r Part_7a}
plot(volume.value, ((1-prop.adults)- (1-prop.infants)), main= "Difference by Volume Value", ylab = "Difference Harvested Proportions", xlab = "Volume", cex = 0.05, col = "dark green")
abline(v = 133.8199, lty = 6)
abline(v = 384.5584, lty = 6)
```

(7)(b) Since curve smoothing is not studied in this course, code is supplied below. Execute the following code to create a smoothed curve to append to the plot in (a). The procedure is to individually smooth (1-prop.adults) and (1-prop.infants) before determining an estimate of the maximum difference. 

```{r Part_7b}

y.loess.a <- loess(1 - prop.adults ~ volume.value, span = 0.25,
	family = c("symmetric"))
y.loess.i <- loess(1 - prop.infants ~ volume.value, span = 0.25,
	family = c("symmetric"))
smooth.difference <- predict(y.loess.a) - predict(y.loess.i)

```

(7)(c) Present a plot of the difference ((1 - prop.adults) - (1 - prop.infants)) versus volume.value with the variable smooth.difference superimposed. Determine the volume.value corresponding to the maximum smoothed difference (Hint:  use *which.max()*). Show the estimated peak location corresponding to the cutoff determined.

Include, side-by-side, the plot from (6)(d) but with a fourth vertical A-B line added. That line should intercept the x-axis at the "max difference" volume determined from the smoothed curve here.

```{r Part_7c, fig.width = 12, fig.height = 6}
cutoff <- volume.value[which.max(smooth.difference)]
cutoff

par(mfrow= c(1,2))

plot(volume.value, ((1-prop.adults) - (1-prop.infants)), main = "Difference by Volume Value", ylab = "Difference Harvested Proportions", xlab = "Volume", cex = 0.05, col = "dark green")
lines(smooth.difference ~ volume.value)
abline(v = cutoff, h = 0.565, lty = 3)
points(cutoff,0.565)
text(cutoff, 0.565, 'Max 262.143', pos = 4, cex = 0.75)


plot(x= volume.value, y=prop.adults, col= "dark green", ylab= "Adult Proportion")
lines(x= volume.value, y= prop.infants, col= "black", xlab= "Volume")
abline(v= median(mydata$VOLUME[mydata$TYPE== "I"]), h= 0.5)
text(133.819, .55, "Median Infant 133.8199")
abline(v= median(mydata$VOLUME[mydata$TYPE== "ADULT"]), h= 0.5)
text(384.5584, .55, "Median Adult 384.5584")
abline(v= max(mydata$VOLUME[mydata$TYPE== "I"]), h= 0.5)
text(526.6836, 0.97, "Max Infant 526.6383")
abline(v = cutoff, h = 0.5, lty = 3)
points(cutoff,0.5)
text(cutoff, 0.5, 'Max Diff 262.143', pos = 4, cex = 0.75)
legend("topright", legend= c("Adult", "Infant"), col= "darkgreen", "black", pch= 1)




```

(7)(d) What separate harvest proportions for infants and adults would result if this cutoff is used? Show the separate harvest proportions. We will actually calculate these proportions in two ways:  first, by 'indexing' and returning the appropriate element of the (1 - prop.adults) and (1 - prop.infants) vectors, and second, by simply counting the number of adults and infants with VOLUME greater than the vlume threshold of interest.

Code for calculating the adult harvest proportion using both approaches is provided.

```{r Part_7d}

sum(mydata[mydata$TYPE == "ADULT", "VOLUME"] >
      volume.value[which.max(smooth.difference)]) / total.adults # [1] 

sum(mydata[mydata$TYPE == "I", "VOLUME"] >
      volume.value[which.max(smooth.difference)]) / total.infants # [1]


```

-----

There are alternative ways to determine cutoffs. Two such cutoffs are described below.

-----

***####  Section 8: (10 points)  ####***

(8)(a) Harvesting of infants in CLASS "A1" must be minimized. The smallest volume.value cutoff that produces a zero harvest of infants from CLASS "A1" may be used as a baseline for comparison with larger cutoffs. Any smaller cutoff would result in harvesting infants from CLASS "A1."  

Compute this cutoff, and the proportions of infants and adults with VOLUME exceeding this cutoff. Code for determining this cutoff is provided. Show these proportions. You may use either the 'indexing' or 'count' approach, or both.

```{r Part_8a}

Alph <- volume.value[volume.value > max(mydata[mydata$CLASS == "A1" & mydata$TYPE == "I", "VOLUME"])][1]
Alph

cutoff2 <- first(volume.value[volume.value > Alph])

cat("Adult Proportion: 0.8260 \n")
cat("Infant Proportion: 0.2872 \n")

```

(8)(b) Next, append one (1) more vertical A-B line to our (6)(d) graph. This time, showing the "zero A1 infants" cutoff from (8)(a). This graph should now have five (5) A-B lines:  "protect all infants," "median infant," "median adult," "max difference" and "zero A1 infants."

```{r Part_8b, fig.width = 12}
plot(x= volume.value, y=prop.adults, col= "dark green", ylab= "Adult Proportion")
lines(x= volume.value, y= prop.infants, col= "black", xlab= "Volume")
abline(v= median(mydata$VOLUME[mydata$TYPE== "I"]), h= 0.5)
text(133.819, .55, "Median Infant 133.8199")
abline(v= median(mydata$VOLUME[mydata$TYPE== "ADULT"]), h= 0.5)
text(384.5584, .55, "Median Adult 384.5584")
abline(v= max(mydata$VOLUME[mydata$TYPE== "I"]), h= 0.5)
text(526.6836, 0.97, "Protect All Infants 526.6383")
abline(v = cutoff, h = 0.5, lty = 3)
points(cutoff,0.5)
text(cutoff, 0.5, 'Max Diff 262.143', pos = 4, cex = 0.75)
abline(v = cutoff2, h = 0.5, lty = 3)
points(cutoff2,0.5)
text(cutoff2, 0.5, 'Zero A1 206.786', pos = 4, cex = 0.75)
legend("topright", legend= c("Adult", "Infant"), col= "darkgreen", "black", pch= 1)


```


***#### Section 9: (5 points) ####***

(9)(a) Construct an ROC curve by plotting (1 - prop.adults) versus (1 - prop.infants). Each point which appears corresponds to a particular volume.value. Show the location of the cutoffs determined in (6), (7) and (8) on this plot and label each. 

```{r Part_9, fig.width = 8.5}
cutoff3<- volume.value[which.min(abs(prop.adults + prop.infants - 1))]
All_Cut<- which(volume.value %in% c(cutoff, cutoff2, cutoff3))

plot((1- prop.infants), (1- prop.adults), main= "ROC Curve: Harvest Proportions", ylab = "Adult Proportions", xlab = "Infant Proportions", cex = 0.5, col = "dark green")
abline(coef = c(0,1), lwd = 2, col= "dark red", lty= 2)
points((1-prop.infants[All_Cut]), (1-prop.adults[All_Cut]))
text(0.287, 0.826, 'Zero A1 206.786')
text(0.176, 0.742, 'Max Diff 262.143')
text(0.0, 0.25, 'Protect All Infants 526.6383')
text(0.498, 0.95, 'Median Infant 133.8199')
text(0.025, 0.5, 'Median Adult 384.5584')



```

(9)(b) Numerically integrate the area under the ROC curve and report your result. This is most easily done with the *auc()* function from the "flux" package.   Areas-under-curve, or AUCs, greater than 0.8 are taken to indicate good discrimination potential. 

```{r Part_9b}
area <- flux::auc(x = (1 - prop.infants), y = (1 - prop.adults))

cat("Area = ", area)

```


***#### Section 10: (10 points) ####***

(10)(a) Prepare a table showing each cutoff along with the following:
 	1) true positive rate (1-prop.adults,
 	2) false positive rate (1-prop.infants),
 	3) harvest proportion of the total population

To calculate the total harvest proportions, you can use the 'count' approach, but ignoring TYPE; simply count the number of individuals (i.e. rows) with VOLUME greater than a given threshold and divide by the total number of individuals in our dataset.
 	
```{r Part_10} 	
table <- matrix(c(262.143, 0.742, 0.176, 0.584,206.786, 0.826, 0.827, 0.676, 133.8214, 0.498, 0.95, 0.5, 384.5584, 0.025, 0.5, 0.5, 526.6383, 0.0, 0.25, 0.03),ncol = 4, byrow = TRUE)

colnames(table) <- c("Cutoff","TPR", "FPR", "Proportional Yield")
rownames(table) <- c("Max Difference", "Zero A1", "Median Infant", "Median Adult", "Protect All Infants")

table<- as.table(table)

table

```
 	
**Essay Question: Based on the ROC curve, it is evident a wide range of possible "cutoffs" exist. Compare and discuss the five cutoffs determined in this assignment.**   

***Answer: (Enter your answer here.)***
The max difference cutoff creates a higher true positive rate with a lower false positive rate, while delivering proportional yield that is lower than the cutoff pertaining to Zero A1 Infants as well as the Median Infants. The Zero A1 Infants cutoff produces the highest true positive rate and while generating the second highest proportional yield rate. the Median Cutoffs are interesting as those yield lowest true positive rates and also produces low false positive rates. It appears the Protect All Infant Cutoff has the 2nd lowest true positive rate and a true positive rate of 0. 



**Final Essay Question:  Assume you are expected to make a presentation of your analysis to the investigators How would you do so?  Consider the following in your answer:**

1. Would you make a specific recommendation or outline various choices and tradeoffs?
2. What qualifications or limitations would you present regarding your analysis?
3. If it is necessary to proceed based on the current analysis, what suggestions would you have for  implementation of a cutoff?  
4. What suggestions would you have for planning future abalone studies of this type? 

***Answer: (Enter your answer here.)***
In this study, we were able to ascertain data to handle both safe and aggressive approaches as it pertains to the harvesting of abalones. I only know what the data tells me, hence there may be some extraneous information that I am unaware of that would effect the decision making process relating to harvesting strategy. The table generated in question 10 would prove to be good information by observing the true postive rate, the false positive rate and the proportional yield. 

However, it is not to say that this information alone should provide a basis for decision making, rather it is a singular piece of information that would aid in the process of determining a harvesting strategy that is best suited for whomever is to make such call. Data manipulation was a large part of this study, hence it would be wise to find more data as it pertains to abalones, possibly more in depth than what was provided here in order to make the best decisions possible. Furthermore, other data sets may yield different results and it is important to compare and contrast said results to determine a best course of action. A basic example of how our data set may differ from other is when a new variable, "ADULT" was created. The determination of what made an abalone and adult, compared to an infant, is unknown for this data set. The same can be said for abalone types when being partitioned by class.

Being that is is absolutely necessary to proceed based on the current analysis, I would suggest a more conservative approach simply due to the ambiguity of the data. If such an approach proves to be successful, then it would be reasonable to incrementally increase harvests. Observing the pros and cons along the way would be key to understanding a proper cutoff. 

I suggest a more systematic approach as it relates to identifying abalone types and classes, which could lead to the mitigation of false positive errors, something that should be desired. Furthermore, the dataset should incorporate affromentioned extraneuos factors to paint a better picture of environmental factors that would effect not only the abalones themsleves but the harvesting process. 