---
title: 'Data Analysis Assignment #1 (50 points total)'
author: "Venit_Michael"
output:
html_document: default
---

```{r setup, include = FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

```


Submit both the .Rmd and .html files for grading. You may remove the instructions and example problem above, but do not remove the YAML metadata block or the first, "setup" code chunk.  Address the steps that appear below and answer all the questions. Be sure to address each question with code and comments as needed.  You may use either base R functions or ggplot2 for the visualizations.

-----

The following code chunk will:

(a) load the "ggplot2", "gridExtra" and "knitr" packages, assuming each has been installed on your machine,
(b) read-in the abalones dataset, defining a new data frame, "mydata,"
(c) return the structure of that data frame, and
(d) calculate new variables, VOLUME and RATIO.

Do not include package installation code in this document. Packages should be installed via the Console or 'Packages' tab. You will also need to download the abalones.csv from the course site to a known location on your machine. Unless a *file.path()* is specified, R will look to directory where this .Rmd is stored when knitting.

```{r analysis_setup1, message = FALSE, warning = FALSE}

# a) Load the ggplot2 and gridExtra packages.

library(ggplot2)
library(gridExtra)
library(knitr)

# b) Use read.csv() to read the abalones.csv into R, assigning the data frame to "mydata."
mydata <- read.csv("~/Desktop/abalones.csv", sep = ",", stringsAsFactors = TRUE)

# c) Use the str() function to verify the structure of "mydata." You should have 1036 observations
# of eight variables.

str(mydata)

# d) Define two new variables, VOLUME and RATIO. Use the following statements to define VOLUME and
# RATIO as variables appended to the data frame "mydata."

mydata$VOLUME <- mydata$LENGTH * mydata$DIAM * mydata$HEIGHT
mydata$RATIO <- mydata$SHUCK / mydata$VOLUME

```

-----

### Test Items starts from here - There are 6 sections - Total 50 points

***##### Section 1: (6 points) Summarizing the data.***

(1)(a) (1 point) Use *summary()* to obtain and present descriptive statistics from mydata.  Use *table()* to present a frequency table using CLASS and RINGS. There should be 115 cells in the table you present.  

```{r Part_1a}
summary(mydata)

classrings<- table(mydata$CLASS, mydata$RINGS)
classrings

```

**Question (1 point):  Briefly discuss the variable types and distributional implications such as potential skewness and outliers.**

***Answer: (Enter your answer here.)***
From the dataset, Abalones seems to be evenly distribuetd across the category of sex. Based on the other variables listed, it appears as if there are outliers for the following: WHIOLE, SHUCK, and RINGS. Such outliers for WHOLE, SHUCK and RINGS appear to be higher values based on locations of the quartiles and the min/max values. This could make the distribution of these categories skewed more to the right. For the category of VOLUME, based on the location of the quartiles, the minimum value seems to be an outlier, which could cause a leftward swekedness if there are more values around the miniumum. 

(1)(b) (1 point) Generate a table of counts using SEX and CLASS. Add margins to this table (Hint: There should be 15 cells in this table plus the marginal totals. Apply *table()* first, then pass the table object to *addmargins()* (Kabacoff Section 7.2 pages 144-147)).  Lastly, present a barplot of these data; ignoring the marginal totals.  

```{r Part_1b}
sexclass<- table(mydata$SEX, mydata$CLASS)
addmargins(sexclass)

ggplot(mydata, aes(CLASS))+ geom_bar(aes(fill= SEX)) +ggtitle("Class by Sex")+xlab("Class")+ ylab("Frequency")

```

**Essay Question (2 points):  Discuss the sex distribution of abalones.  What stands out about the distribution of abalones by CLASS?**

As it pertains to the distribution of classes by Class by Sex, infants dominate classes A1 and A2 and are skewed right compared to Male and Females. Male abaolnes appear to be most frequent in class A3. There seems to be a rather consistent ratio between Male and Female Abalones consistently across each class.  


(1)(c) (1 point) Select a simple random sample of 200 observations from "mydata" and identify this sample as "work." Use *set.seed(123)* prior to drawing this sample. Do not change the number 123. Note that *sample()* "takes a sample of the specified size from the elements of x." We cannot sample directly from "mydata." Instead, we need to sample from the integers, 1 to 1036, representing the rows of "mydata." Then, select those rows from the data frame (Kabacoff Section 4.10.5 page 87). 

Using "work", construct a scatterplot matrix of variables 2-6 with *plot(work[, 2:6])* (these are the continuous variables excluding VOLUME and RATIO). The sample "work" will not be used in the remainder of the assignment. 

```{r Part_1c}
set.seed(123)
work<- mydata[sample(1:nrow(mydata), 200, replace = FALSE),]
plot(work[, 2:6])
```

-----

***##### Section 2: (5 points) Summarizing the data using graphics.***

(2)(a) (1 point) Use "mydata" to plot WHOLE versus VOLUME. Color code data points by CLASS.

```{r Part_2a}
ggplot(mydata, aes(x= WHOLE, y= VOLUME, color= CLASS))+ geom_point()+ ggtitle("Whole v. Volume")+ xlab("Whole (grams)")+ ylab(expression("Volume (cm)"^"3",")"))

```


(2)(b) (2 points) Use "mydata" to plot SHUCK versus WHOLE with WHOLE on the horizontal axis. Color code data points by CLASS. As an aid to interpretation, determine the maximum value of the ratio of SHUCK to WHOLE.  Add to the chart a straight line with zero intercept using this maximum value as the slope of the line. If you are using the 'base R' *plot()* function, you may use *abline()* to add this line to the plot. Use *help(abline)* in R to determine the coding for the slope and intercept arguments in the functions. If you are using ggplot2 for visualizations, *geom_abline()* should be used. 

```{r Part_2b}
ggplot(mydata, aes(x= WHOLE, y= SHUCK, color= CLASS))+ geom_point()+ geom_abline(mydata, intercept = 0, slope= max(mydata$SHUCK/mydata$WHOLE))+ggtitle("Shuck v. Whole")+ xlab("Whole (grams)")+ylab("Shuck (grams)")

```

**Essay Question (2 points):  How does the variability in this plot differ from the plot in (a)?  Compare the two displays.  Keep in mind that SHUCK is a part of WHOLE.  Consider the location of the different age classes.**

***Answer: (Enter your answer here.)***
In plot 1, the distribution seems to be linear, but the data points are less consistent than that of plot 2. Also, for plot 2 (Shuck v. Whole), many of the data points lie below the slope of the line, indicating that abalones have a lower whole weight while mainitaing higher shuck. It appears as though abaloines from class A5 have lower shuck which affects their whole weight. 

-----

***### Section 3: (8 points) Getting insights about the data using graphs.***

(3)(a) (2 points) Use "mydata" to create a multi-figured plot with histograms, boxplots and Q-Q plots of RATIO differentiated by sex. This can be done using *par(mfrow = c(3,3))* and base R or *grid.arrange()* and ggplot2. The first row would show the histograms, the second row the boxplots and the third row the Q-Q plots. Be sure these displays are legible.  

```{r Part_3a}
par(mfrow = c(3,3))

hist(mydata$RATIO[mydata$SEX=="I"], col = "purple", main = "Abalone Ratio (Infants)", xlab = "")
hist(mydata$RATIO[mydata$SEX=="F"], col = "green", main = "Abalone Ratio (Females)", xlab = "")
hist(mydata$RATIO[mydata$SEX=="M"], col = "blue", main = "Abalone Ratio (Males)", xlab = "")

boxplot(mydata$RATIO[mydata$SEX=="I"], col = "purple", main = "Abalone Ratio (Infants)", xlab = "")
boxplot(mydata$RATIO[mydata$SEX=="F"], col = "green", main = "Abalone Ratio (Females)", xlab = "")
boxplot(mydata$RATIO[mydata$SEX=="M"], col = "blue", main = "Abalone Ratio (Males)", xlab = "")

qqnorm(mydata$RATIO[mydata$SEX=="I"], col = "purple", main = "Abalone Ratio (Infants)")
qqline(mydata$RATIO[mydata$SEX=="I"])
qqnorm(mydata$RATIO[mydata$SEX=="F"], col = "green", main = "Abalone Ratio (Females)")
qqline(mydata$RATIO[mydata$SEX=="F"])
qqnorm(mydata$RATIO[mydata$SEX=="M"], col = "blue", main = "Abalone Ratio (Males)")
qqline(mydata$RATIO[mydata$SEX=="M"])
```

**Essay Question (2 points):  How does the variability in this plot differ from the plot in (a)?  Compare the two displays.  Keep in mind that SHUCK is a part of WHOLE.  Consider the location of the different age classes.**

**Answer:** 
When the data is separated based on Sex, each plot contains outliers on the upper righthand side. Each plot appears normally distributed with a skew to the right. the 3 QQ plots exhibit more curvature at the end of the distribution, which would indicate that these are non-normal distributions.

(3)(b) (2 points) The boxplots in (3)(a) indicate that there are outlying RATIOs for each sex. *boxplot.stats()* can be used to identify outlying values of a vector. Present the abalones with these outlying RATIO values along with their associated variables in "mydata". Display the observations by passing a data frame to the kable() function. Basically, we want to output those rows of "mydata" with an outlying RATIO, but we want to determine outliers looking separately at infants, females and males.

```{r Part_3b}

outliers<- boxplot.stats(mydata$RATIO, coef = 1.5)$out
index<- match(outliers, mydata$RATIO)
ID<- seq_len(nrow(mydata))
mydataindexed<- cbind(ID, mydata)

print(mydataindexed[index,])


```

**Essay Question (2 points):  What are your observations regarding the results in (3)(b)?**

***Answer: (Enter your answer here.)***
Out of the sexes, infants have the highest amount of outliers. The majority of outliers can be found mostly in classes A1, A2, and A3. Male abalone outliers lie in mostly in class A3, while outliers for infant abalones lie in class A1.
-----

***### Section 4: (8 points) Getting insights about possible predictors.***

(4)(a) (3 points) With "mydata," display side-by-side boxplots for VOLUME and WHOLE, each differentiated by CLASS There should be five boxes for VOLUME and five for WHOLE. Also, display side-by-side scatterplots:  VOLUME and WHOLE versus RINGS. Present these four figures in one graphic:  the boxplots in one row and the scatterplots in a second row. Base R or ggplot2 may be used.


```{r Part_4a}
par(mfrow = c(2,2))

boxplot(mydata$VOLUME~mydata$CLASS, col = "lightblue", main = "Volume  by Class", xlab = "", ylab = "Volume")
boxplot(mydata$WHOLE~mydata$CLASS, col = "lightgreen", main = "Whole Weight by Class", xlab = "", ylab = "Whole Weight")

plot(mydata$RINGS~mydata$VOLUME, main = "Volume v. Rings", xlab = "Rings", ylab = "Volume")
plot(mydata$RINGS~mydata$WHOLE, main = "Whole Weight v. Rings", xlab = "Rings", ylab = "Whole Weight")

```

**Essay Question (5 points) How well do you think these variables would perform as predictors of age?  Explain.**

***Answer: (Enter your answer here.)***
From the plots above, we notice large variability when observing volume versus rings as well as whole wieght versus rings. However, these plots do show a positive (upward) trend, which would indicate that Rings, Volume and Whole Wieght would be good indicators of age. It is a certainty that volume and whole weight increases and abalones receive more rings. 

-----

***### Section 5: (12 points) Getting insights regarding different groups in the data.***

(5)(a) (2 points) Use *aggregate()* with "mydata" to compute the mean values of VOLUME, SHUCK and RATIO for each combination of SEX and CLASS. Then, using *matrix()*, create matrices of the mean values. Using the "dimnames" argument within *matrix()* or the *rownames()* and *colnames()* functions on the matrices, label the rows by SEX and columns by CLASS. Present the three matrices (Kabacoff Section 5.6.2, p. 110-111). The *kable()* function is useful for this purpose.  You do not need to be concerned with the number of digits presented.

```{r Part_5a}
shuck <- aggregate(SHUCK~CLASS+SEX, data = mydata, mean)
volume <- aggregate(VOLUME~CLASS+SEX, data = mydata, mean)
rat <- aggregate(RATIO~CLASS+SEX, data = mydata, mean)

shuckmat <- matrix(data = shuck[,3], nrow=3, byrow = FALSE)
colnames(shuckmat) <- c("A1","A2","A3","A4","A5")
rownames(shuckmat) <- c("Shuck F:", "Shuck I:", "Shuck M:")

volumemat <- matrix(data = volume[,3], nrow=3, byrow = FALSE)
colnames(volumemat) <- c("A1","A2","A3","A4","A5")
rownames(volumemat) <- c("Volume F:", "Volume I:", "Volume M:")

ratiomat <- matrix(data = rat[,3], nrow=3, byrow = FALSE)
colnames(ratiomat) <- c("A1","A2","A3","A4","A5")
rownames(ratiomat) <- c("Ratio F:", "Ratio I:", "Ratio M:")

shuckmat
volumemat
ratiomat

```

(5)(b) (3 points) Present three graphs. Each graph should include three lines, one for each sex. The first should show mean RATIO versus CLASS; the second, mean VOLUME versus CLASS; the third, mean SHUCK versus CLASS. This may be done with the 'base R' *interaction.plot()* function or with ggplot2 using *grid.arrange()*.

```{r Part_5b, fig.width = 9}
par(mfrow= c(3,1))

interaction.plot(response = mydata$RATIO, trace.factor = mydata$SEX, x.factor = mydata$CLASS, fun = mean, trace.label = "Sex", xlab= "Class", ylab = "Mean Ratio")

interaction.plot(response = mydata$VOLUME, trace.factor = mydata$SEX, x.factor = mydata$CLASS, fun = mean, trace.label = "Sex", xlab = "Class", ylab = "Mean Volume")

interaction.plot(response = mydata$SHUCK, trace.factor = mydata$SEX, x.factor = mydata$CLASS, fun = mean, trace.label = "Sex", xlab = "Class", ylab = "Mean Shuck")

```

**Essay Question (2 points):  What questions do these plots raise?  Consider aging and sex differences.**

***Answer: (Enter your answer here.)***
The graph in which mean ratio is compared by class indicates a decreasing ratio as class increases for all 3 sexes. However, this contrasts what is being seen with mean shuck, particularly with clsses A4 and A5. For mean volume and mean shuck, there appears to be a point in which abalones plateau as can be seen for classes A4 and A5 in the instances. This raises the question of whether or not an abalone is worth eating at class 2 due to the high mean ratio? Or are the better to consume in class 4 when they have their highest Shuck ratio?



5(c) (3 points) Present four boxplots using *par(mfrow = c(2, 2)* or *grid.arrange()*. The first line should show VOLUME by RINGS for the infants and, separately, for the adult; factor levels "M" and "F," combined. The second line should show WHOLE by RINGS for the infants and, separately, for the adults. Since the data are sparse beyond 15 rings, limit the displays to less than 16 rings. One way to accomplish this is to generate a new data set using subset() to select RINGS < 16.  Use ylim = c(0, 1100) for VOLUME and ylim = c(0, 400) for WHOLE.  If you wish to reorder the displays for presentation purposes or use ggplot2 go ahead.

```{r Part_5c}
par(mfrow = c(2,2))
limited <- subset(mydata, mydata$RINGS < 16)

boxplot(limited$VOLUME[limited$SEX == "I"] ~ limited$RINGS[limited$SEX == "I"], ylim=c(0,1100), main= "Infant Volume v. Rings", col= "blue", ylab = "Volume", xlab = "Rings")
boxplot(limited$VOLUME[limited$SEX != "I"] ~ limited$RINGS[limited$SEX != "I"], ylim=c(0,1100), main= "Adult Volume v. Rings", col= "green", ylab = "Volume", xlab = "Rings")
boxplot(limited$WHOLE[limited$SEX == "I"] ~ limited$RINGS[limited$SEX == "I"], ylim=c(0,400), main= "Infant Whole Weight v. Rings", col= "blue", ylab = "Whole Weight", xlab = "Rings")
boxplot(limited$WHOLE[limited$SEX != "I"] ~ limited$RINGS[limited$SEX != "I"], ylim=c(0,400), main= "Adult Whole Weight v. Rings", col= "green", ylab = "Whole Weight", xlab = "Rings")  

```

**Essay Question (2 points):  What do these displays suggest about abalone growth?  Also, compare the infant and adult displays.  What differences stand out?**

***Answer: (Enter your answer here.)*** 
The graphs above show that the growth of adult abalones vary more than that of infant abalones. The whiskers in the graphs for adult abalones indicate higher dispersion as the number of rings grow. As mentioned previously, it appears that volume and whole weight increase with ring count. 

-----

***### Section 6: (11 points) Conclusions from the Exploratory Data Analysis (EDA).***


**Conclusions**


**Essay Question 1) (5 points)	Based solely on these data, what are plausible statistical reasons that explain the failure of the original study? Consider to what extent physical measurements may be used for age prediction.**

***Answer: (Enter your answer here.)***
There were multiple outliers in the data set which were not accounted for and fixed beforehand. The graphics rendered in this exam appear to be skewed to the right. Ring count seems to be the most indicative of age and is a good qualitative measure as can be seen when comparing rings to volume and whole weight. As it pertains to to abalone sex, it was a surprise to see infants in all 5 classes. This raises questions regarding the data set as the there is uncertainty in regard to the classification of abalones as it relates to this particular set of data. There lies the possibility that error and bias played a part in the organization of the data.  

**Essay Question 2) (3 points) Do not refer to the abalone data or study.  If you were presented with an overall histogram and summary statistics from a sample of some population or phenomenon and no other information, what questions might you ask before accepting them as representative of the sampled population or phenomenon?**

***Answer: (Enter your answer here.)***
I would ask about the environment in which the study was taken as the obsevred data could be differentiated by environmental facotrs. I believe it is important to gain context on the study at hand. Additionally, I would want to know about the kind of distribution for the sample/phenomenon being observed and if the sample taken in representative of the overall population. It would also be helpful to observe the outliers in the data set and if there are any external factors contributing to the recording of said outliers. 

**Essay Question 3)  (3 points)	Do not refer to the abalone data or study.  What do you see as difficulties analyzing data derived from observational studies? Can causality be determined?  What might be learned from such studies?**

***Answer: (Enter your answer here.)***
Data that is observed from observational studies can sometimes be subject to bias and error. Bias and error are difficult to account for and can affect an entire data set and how it was recorded, which can result in misleading cocnlusions. I believe causality would not be able to be determined as we do not have any context for the study at hand. It would be much easier to determine correlation between variables rather than cause. I believe it is important to sample properly so that the data reflects that of the entire population, this way accurate conclusions can be made. 
